Word Learning as Network Growth: a Cross-linguistic Analysis

### 2018-01-25 cogsci/cogsci.tex
changes in CogSci manuscript

\begin{abstract}
Children tend to learn first the words that are highly connected in the
lexical network (i.e., words that are connected to a variety of other
words). The process by which connectivity in both the phonological and
semantic netwoks come to influence word learning is still not very well
understood. One view suggests that learning is driven primarily by
\emph{internal} connectivity in the child's available lexicon. Another
view suggests that \emph{external} connectivity in the learning
environment is what makes words salient and readily learned. The present
study tests both scenarios systematically in both the phonological and
the the semantic domains, and across 8 languages. We generate the
results obtained by Hills et al. (2009) with English semantic networks,
and show that external connectivity in the learning environment, rather
than internal connectivity, drives growth of both semantic and
phonological networks. Finally, we demonstrate that these effects hold
cross-linguistically.

\textbf{Keywords:}
semantic network, phonological network, word learning
\end{abstract}

@@

What factors shape vocabulary learning over the course of early
childhood? To investigate this question, scientists have adopted
multiple research strategies, from controlled laboratory experiments to
corpus analysis. One strategy consists in documenting the timeline of
words' acquisition, and studying the properies that make words easy or
hard to learn. For example, within a lexical category, words that are
more frequent in child-directed speech are acquired earlier (J. C.
Goodman, Dale, \& Li, 2008). Other factors include word length, the mean
length of utterances in which the word occurs, and how concrete the word
is (see Braginsky, Yurovsky, Marchman, \& Frank, 2016).

@@

Studies that investigted lexical network growth focused on semantic
networks using English data only (Hills et al., 2010, 2009; Steyvers \&
Tenenbaum, 2005). The novelty of the current study is that it
investigates whether phonological networks, like semantic networks, grow
by PAC, or if they rather grow by PAT. It also provides a systematic
comparision of network growth scenarios in both the phonological and the
semantic domains and assesses their relative contribution to the
learning process. Moreover, it tests the generality of the findings
across eight languages.

The paper is organized as follows. First, we describe the dataset we
used, the procedure we followed to construct both semantic and
phonological networks. Next, we show the results of the various analyses
we conducted comparing network growth scenarios across languages. Then
we test the overall contribution of the networks in the learning process
relative to other known predictors of age of acquisition (frequency and
length). Finally we discuss the major results in relation to known
experimental facts in language development.

@@

@@

We used data from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman,
2017), an open repository aggregating cross-linguistic language
developmental data of the MacArthur-Bates Communicative Development
Inventory (CDI), a parent report vocabulary checklist. We used the
\emph{Words and Sentence} version of the CDI which contains the
productive vocabulary of toddlers (age varied between 16 to 36 months).
Following previous studies (Hills et al., 2009; e.g., Storkel, 2009), we
restricted our analysis to the category of nouns, since there are known
discrepancies between the learning of nouns, predicates, and function
words (e.g., Braginsky et al., 2016). The age of acquisition was defined
by the month at which a word was produced by at least 50\% of children
(J. C. Goodman et al., 2008).

@@

We used as an index of semantic relatedness the Florida Free Association
Norms (Nelson, McEvoy, \& Schreiber, 1998). This dataset was collected
by giving adult participants a word (the cue), and asking them to write
the first word that comes to mind (the target). For example, when given
the word ``ball'', they might answer with the word ``game''. A pair of
nodes were connected by a directed link from the cue to the target if
there was a cue-target relationship between these nodes in the
association norms. Each node was characterized by its ``indegree'',
which represents the number of links for which the word was the target.
Following Hills et al. (2009), we used as a proxy for the learning
environment, the network constructed using the full set of nouns in the
CDI data in each language (e.g., in English this corresponds to the
network by 30 months).

@@

We used as a measure of phonological relatedness between two nodes the
Levenshtein (edit) distance between their phonological forms. The
measure counts the minimum number of operations (insertions, deletions,
substitutions) required to change one string into another. We generated
approximate International Phonetic Alphabet (IPA) trnascriptions from
the orthographic transcription, across languages, using the open source
text-to-speech software
\textbf{\href{http://http://espeak.sourceforge.net/}{Espeak}.}

@@

We start by examining how connectivity in the global netowrk influences
the age of acqusition. Regardless of the growth mechanism that might
have given rise to this netowrk (i.e., PAC or PAT), we expect nodes with
the higher degrees in this network to be learned earlier than nodes with
the lower degrees. The network was constructed using nouns at the oldest
age for which we have CDI data. Figure \ref{fig:data_all} shows how the
age of acquisition for each word varies as a function of its degree (or
indegree for the semantic network). For ease of visual comparison, the
predictor (i.e., the degree) was centered and scaled across languages.
The plots show, overall, a negative correlation between the month of
acquisition and the degree, indicating that nouns with higher degrees
are generally learned earlier.

@@

Where \(\beta\) is the fitting parameter. A positive value of \(\beta\)
means that words with higher growth values \(d_i\) are acquired first,
and a negative value means that words with lower growth values are
acquired first. As explained in Figure \ref{fig:growth}, each candidate
word \(w_i\) is characterized with a growth values \(d_i\) that depends
on the growth scenario tested. The normalization includes all words that
could be learned at that month.

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilisitic programming language
WebPPL (N. Goodman \& Stuhlmüller, 2014). We started with a uniform
distribution, and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been actually learned at that month (using formula
1). We ended up with a posterior distribution of \(\beta\), which we
summarized in Figure \ref{fig:pred_ind}.

@@

We saw that the way semantic and phonological information is structured
in the learning environment (i.e., PAC) contribute to noun learning
across languages. However, we know that other factors influence learning
as well (e.g., Braginsky et al., 2016). In what follows, we will
investigate how semantic and phonological connectivity interact with two
other factors. The first one is word frequency, a well studied factor
shown to predict the age of acquisition in a reliable fashion (e.g. J.
C. Goodman et al., 2008). The second factor is word length, which
correlates with phonological connectivity (Pisoni, Nusbaum, Luce, \&
Slowiaczek, 1985).

@@

The present study provided a comprehensive analysis of how lexical
connectivity influences the age of acquisition of nouns in toddlers. We
compared two network growth scenarios and assessed the relative
contribution of phonological and semantic information in 8 languages.
Part of the findings largely replicate the results obtained in Hills et
al. (2009), i.e., semantic networks (based on free associations) grow by
preferential acquisition, not by preferential attachment. Another
finding was that phonological networks also grow primarily by
preferential acquisition, especially when both scenarios (PAT and PAC)
were pitted against each other in the same model. These findings
generalize well across languages. Moreoever, both semantic and
phonological connectivity in the learning environment (i.e., PAC)
predict growth in a consistent way across many languages. However, when
pitted against other known predictors of age of acquisition (word
frequency and length), the effect of word connectivity shows a
cross-linguistic variation, predicting learning in some languages, but
not in others. Despite this cross-lingusitic variation, both
phonological and semantic connectivity contribute to the overall
learning (when data is pooled across languages).

@@

A major results of the study is that children start by learning words
that have high phonological and semantic similarity with a variety of
other words in the learning environment, not in the child's available
lexicon. This suggests that children are sensitive to connectivity even
without having first acquired the connected words. How can children
indirectly detect highly connected words, and why would such words be
more readily learned? In the semantic case, free association can be
predicted through the patterns of word co-occurrence (Griffiths,
Steyvers, \& Tenenbaum, 2007), meaning that highly connected words tend
to be the words that co-occur with many other words in various contexts.
One possibility, suggested by Hills et al. (2010), is that the referents
of such words are more easily disambiguated from other potential
referents. Such effect was demonstrated by cross-situational learning
experiments (Smith \& Yu, 2008). In the phonological case, connectivity
is inherently correlated with phonotactic probability (Vitevitch, Luce,
Pisoni, \& Auer, 1999). That is, highly connected words tend to be made
of frequent sound sequences. We know even infant (whose vocabulary is
still very rudimentary) develop a sensitivity for high frequency sound
sequences in the ambient language (Jusczyk, Luce, \& Charles-Luce,
1994). Interestingly, it was shown that phonotactic probability
facilitates learning and recognition of novel words in toddlers and
preschoolers (MacRoy-Higgins, Shafer, Schwartz, \& Marton, 2014;
Storkel, 2001). Thus, children's ability to keep track of co-occurrence
statistics (both at the semantic and phonological level) might explain
their sensitivity and preference for highly connected words in the
learning envirnoment.

Finally, this study shares a number of limitations with previous studies
using similar datasets. For instance, we used normative age of
acquisition. However, individual children may follow different learning
trajectories. The use of longitudinal data would allow us to assess how
the aggregate behavior relates to these individual trajectories (e.g.,
typical vs.~late talkers, Beckage et al., 2011). Moreover, the results
obtained in this study provide correlational but not causal evidence.
Thus, conclusions drawn from our network analysis would require parallel
evidence, especially from controlled behavioral experiments.

@@

\hypertarget{ref-pisoni1985}{}
Pisoni, D. B., Nusbaum, H. C., Luce, P. A., \& Slowiaczek, L. M. (1985).
Speech perception, word recognition and the structure of the lexicon.
\emph{Speech Communication}, \emph{4}(1), 75--95.

>>>>

\begin{abstract}
Children tend to produce first the words that are connected to a variety
of other words along both the phonological and the semantic dimensions.
Though this connectivity effect has been extensively documented, little
is known about the underlying developmental mechanism. One view suggests
that learning is primarily driven by a network growth model where highly
connected words in the child's ealy lexicon attract similar words.
Another view suggests that learning is driven, not by highly connected
words in the early internal lexicon, but by highly connected word in the
external learning environment. The present study tests both scenarios
systematically in both the phonological and the semantic domains, and
across 8 languages. We show that external connectivity in the learning
environment drives growth in both the semantic and the phonological
networks, and that this pattern is consistent cross-linguistically. The
findings suggest a word learning mechanism where children harness their
statistical learning abilities to (indirectly) detect and learn highly
connected words in the learning environment.

\textbf{Keywords:}
semantic network, phonological network, network growth, mechanism of
word learning
\end{abstract}

@@

What factors shape vocabulary learning over the course of early
childhood? To investigate this question, scientists have adopted
multiple research strategies, from controlled laboratory experiments to
corpus analysis. One strategy consists in documenting the timeline of
words' acquisition, and studying the properties that make words easy or
hard to learn. For example, within a lexical category, words that are
more frequent in child-directed speech are acquired earlier (J. C.
Goodman, Dale, \& Li, 2008). Other factors include word length, the mean
length of utterances in which the word occurs, and how concrete the word
is (see Braginsky, Yurovsky, Marchman, \& Frank, 2016).

@@

Studies that investigated lexical network growth focused on semantic
networks using English data only (Hills et al., 2010, 2009; Steyvers \&
Tenenbaum, 2005). The novelty of the current study is that it
investigates whether phonological networks, like semantic networks, grow
by PAC, or if they rather grow by PAT. It also provides a systematic
comparison of network growth scenarios in the phonological and the
semantic domains and assesses their relative contribution to the
learning process. Moreover, it tests the generality of the findings
across eight languages.

The paper is organized as follows. First, we describe the dataset we
used, the procedure we followed to construct both semantic and
phonological networks. Next, we show the results of the various analyses
we conducted comparing network growth scenarios across languages. Then
we test the overall contribution of the networks in the learning process
relative to other known predictors of age of acquisition (frequency and
length). Finally we discuss the major results in relation to known
experimental facts in language development, and we speculate about the
potential learning mechanism that might drive early vocabulary learning.

@@

{\centering \includegraphics{figs/degree_distribution-1}

\caption{\label{fig:degree_distribution}Log-log plot of the cumulative degree distribution function for the global phonological and semantic networks across languages. A perfect power law distribution should appear as a straight line in this graph.}\label{fig:degree_distribution}
\end{figure*}
\end{CodeChunk}

@@

We used data from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman,
2017), an open repository aggregating cross-linguistic language
developmental data of the MacArthur-Bates Communicative Development
Inventory (CDI), a parent report vocabulary checklist. We used the
\emph{Words and Sentence} version of the CDI which contains the
productive vocabulary of toddlers (age varied between 16 to 36 months).
Following previous studies (Hills et al., 2009; e.g., Storkel, 2009), we
restricted our analysis to the category of nouns. The age of acquisition
was defined by the month at which a word was produced by at least 50\%
of children (J. C. Goodman et al., 2008).

@@

We used as an index of semantic relatedness the Florida Free Association
Norms (Nelson, McEvoy, \& Schreiber, 1998). This dataset was collected
by giving adult participants a word (the cue), and asking them to write
the first word that comes to mind (the target). For example, when given
the word ``ball'', they might answer with the word ``game''. A pair of
nodes were connected by a directed link from the cue to the target if
there was a cue-target relationship between these nodes in the
association norms. Each node was characterized by its ``indegree'',
which represents the number of links for which the word was the target.
To model growth from month to month, we constructed a different network
at each month, made of words that have been learned at that month.
Following Hills et al. (2009), we used as a proxy for the learning
environment, the network constructed using the full set of nouns in the
CDI data in each language (e.g., in English this corresponds to the
network by 30 months).

@@

We used as a measure of phonological relatedness between two nodes the
Levenshtein (edit) distance between their phonological forms. The
measure counts the minimum number of operations (insertions, deletions,
substitutions) required to change one string into another. We generated
approximate International Phonetic Alphabet (IPA) transcriptions from
the orthographic transcription, across languages, using the open source
text-to-speech software
\textbf{\href{http://http://espeak.sourceforge.net/}{Espeak}.}

@@

\subsection{Preliminary analysis}\label{preliminary-analysis}

We start by examining how connectivity in the global network influences
the age of acquisition. Regardless of the growth mechanism that might
have given rise to this network (i.e., PAC or PAT), we expect nodes with
the higher degrees in this network to be learned earlier than nodes with
the lower degrees. The global network was constructed using nouns at the
oldest age for which we have CDI data. Figure \ref{fig:data_all} shows
how the age of acquisition for each word varies as a function of its
degree (or indegree for the semantic network). For ease of visual
comparison, the predictor (i.e., the degree) was centered and scaled
across languages. The plots show, overall, a negative correlation
between the month of acquisition and the degree, indicating that nouns
with higher degrees are generally learned earlier.

We also analysed the global network degree distribution. PAT is known to
generate networks with a power-law degree distribution (Barabasi \&
Albert, 1999). Thus, if the networks display this property, it means
that PAT cannot be ruled out \emph{a priori} as a potential growth
mechanism. In contrast, if there is strong evidence against a power law
distribution, then PAT (but not PAC) can be ruled out. The distributions
are shown in Figure \ref{fig:degree_distribution}. Overall, the results
do not provide strong evidence against a power-law distribution.
Moreover, the degree to which the distribution approximates a power law
varies across dimension and languages. For a more conclusive test, we
fit explicit growth models to the data (next section).

@@

Where \(\beta\) is the fitting parameter. A positive value of \(\beta\)
means that words with higher growth values \(d_i\) are acquired first,
and a negative value means that words with lower growth values are
acquired first. As explained in Figure \ref{fig:growth}, each candidate
word \(w_i\) is characterized with a growth values \(d_i\) that depends
on the growth scenario tested. In PAT, the growth value of a given word
is equal to the average degree of the existing nodes that it would
attach to. In PAC, the growth value of a word is equal to its degree in
the entire network. The normalization includes all words that could be
learned at that month.

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilistic programming language
WebPPL (N. Goodman \& Stuhlmüller, 2014). We started with a uniform
distribution, and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been actually learned at that month (using formula
1). We ended up with a posterior distribution of \(\beta\), which we
summarized in Figure \ref{fig:pred_ind}.

@@

We saw that the way semantic and phonological information is structured
in the learning environment (i.e., PAC) contribute to noun learning
across languages. However, we know that other factors influence learning
as well (e.g., Braginsky et al., 2016). In what follows, we will
investigate how semantic and phonological connectivity interact with two
other factors. The first one is word frequency, a well studied factor
shown to predict the age of acquisition in a reliable fashion (e.g. J.
C. Goodman et al., 2008). The second factor is word length, which
correlates with phonological connectivity.

@@

The present study provided a comprehensive analysis of how lexical
connectivity influences the age of acquisition of nouns in toddlers. We
compared two network growth scenarios and assessed the relative
contribution of phonological and semantic information in 8 languages.
Part of the findings largely replicate the results obtained in Hills et
al. (2009), i.e., semantic networks (based on free associations) grow by
preferential acquisition, not by preferential attachment. Another
finding was that phonological networks also grow primarily by
preferential acquisition, especially when both scenarios (PAT and PAC)
were pitted against each other in the same model. These findings
generalize well across languages. Moreover, both semantic and
phonological connectivity in the learning environment (i.e., PAC)
predict growth in a consistent way across many languages. However, when
pitted against other known predictors of age of acquisition (word
frequency and length), the effect of word connectivity shows a
cross-linguistic variation, predicting learning in some languages, but
not in others. Despite this cross-linguistic variation, both
phonological and semantic connectivity contribute to the overall
learning (when data is pooled across languages).

@@

A major results of the study is that children start by learning words
that have high phonological and semantic similarity with a variety of
other words in the learning environment, not in the child's available
lexicon. This suggests that children are sensitive to connectivity even
without having first acquired the connected words. How can children
indirectly detect highly connected words, and why would such words be
more readily learned? In the semantic case, free association can be
predicted through the patterns of word co-occurrence (Griffiths,
Steyvers, \& Tenenbaum, 2007), meaning that highly connected words tend
to be the words that co-occur with many other words in various contexts.
One possibility, suggested by Hills et al. (2010), is that the referents
of such words are more easily disambiguated from other potential
referents. Such effect was demonstrated by cross-situational learning
experiments (Smith \& Yu, 2008).

In the phonological case, connectivity is inherently correlated with
phonotactic probability (Vitevitch, Luce, Pisoni, \& Auer, 1999). That
is, highly connected words tend to be made of frequent sound sequences.
We know even infant (whose vocabulary is still very rudimentary) develop
a sensitivity for high frequency sound sequences in the ambient language
(Jusczyk, Luce, \& Charles-Luce, 1994). Interestingly, it was shown that
phonotactic probability facilitates learning and recognition of novel
words in toddlers and preschoolers (MacRoy-Higgins, Shafer, Schwartz, \&
Marton, 2014; Storkel, 2001). Thus, children's ability to keep track of
co-occurrence statistics (both at the semantic and phonological level)
might explain their sensitivity and preference for highly connected
words in the learning environment.

Finally, this study shares a number of limitations with previous studies
using similar datasets. In particular, the results provide correlational
but not causal evidence. Thus, the conclusions of this study require
parallel evidence, especially from controlled behavioral experiments.

@@
